---
title: "Hands-on Exercise 4(1)"
author: "farrahmf"
---

## Visual Statistical Analysis

### Loading packages and importing data

```{r}
pacman::p_load(ggstatsplot, rstantools, PMCMRplus, tidyverse)

exam_data <- read_csv("data/Exam_data.csv")
```

### One-sample test

Using **gghistostats()**.

```{r}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

### Two-sample mean test

Using **ggbetweenstats()**.

```{r}
ggbetweenstats(
  data = exam_data,
  x = GENDER,
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

### Oneway ANOVA test

Using **ggbetweenstats()**.

```{r}
ggbetweenstats(
  data = exam_data,
  x = RACE,
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE,
  pairwise.comparisons = TRUE,
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

### Significant Test of Correlation

Using **ggscatterscats()**.

```{r}
ggscatterstats(
  data = exam_data,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE
)
```

### Significant Test of Association (Dependence)

Using **ggbarstats()**.

```{r}
exam1 <- exam_data %>%
  mutate(MATHS_bins =
           cut(MATHS,
               breaks = c(0, 60, 75, 85, 100)))

ggbarstats(exam1,
           x = MATHS_bins,
           y = GENDER)
```

## Visualising Models

### Loading packages and installing data

```{r}
pacman::p_load(readxl, performance, parameters, see)

car_resale <- read_xls("data/ToyotaCorolla.xls", "data")

car_resale
```

### Multiple Regression Model

Using **lm()** from Base Stats of R.

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + Weight + Guarantee_Period,
            data = car_resale)

model
```

### Model Diagnostic: Checking for multicollinearity

Using **check_collinearity()** from the **performance** package.

```{r}
check_collinearity(model)
```

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

### Model Diagnostic: Checking normality assumption

Using **check_normality()** from the **performance** package.

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + Weight + Guarantee_Period,
             data = car_resale)

check_n <- check_normality(model1)

plot(check_n)
```

### Model Diagnostic: Checking for heteroscedasticity

Using **check_heteroscedasticity()** from the **performance** package.

```{r}
check_h <- check_heteroscedasticity(model1)

plot(check_h)
```

### Model Diagnostic: Complete check

Using **check_model()**.

```{r}
check_model(model1)
```

### Visualising Regression Parameters

Using **plot()** of **see** package and **parameters()** of **parameters** package to visualise the parameters of a regression model.

```{r}
plot(parameters(model1))
```

Using **ggcoefstats()** of **ggstatsplot** package to visualise the parameters of a regression model.

```{r}
ggcoefstats(model1, output = "plot")
```
