---
title: "Geospatial Point Data"
author: "farrahmf"
date: "25 February 2023"
date-modified: "`r Sys.Date()`"
execute: 
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

### Loading packages and importing dataset

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

```{r}
sgpools <- read_csv("aspatial/SGPools_svy21.csv")
list(sgpools)
```

## Creating an sf data frame from an aspatial data frame

*\
According to [Wikipedia](https://en.wikipedia.org/wiki/Simple_Features), Simple Features (officially Simple Feature Access) is a set of standards that specify a common storage and access model of geographic feature made of mostly two-dimensional geometries (point, line, polygon, multi-point, multi-line, etc.) used by geographic information systems. It is formalized by both the Open Geospatial Consortium (OGC) and the International Organization for Standardization (ISO).*

Convert **sgpools** into a simple feature data frame. A new column called *geometry* will be added, which has the coordinates in the specified [European Petroleum Survey (EPSG)](https://en.wikipedia.org/wiki/EPSG_Geodetic_Parameter_Dataset) format.

```{r}
sgpools_sf <- st_as_sf(sgpools, 
                       coords = c("XCOORD", "YCOORD"),
                       crs= 3414)
```

*(Note: The* crs *argument requires the coordinates system in [epsg](https://epsg.io/) format.)*

## Drawing Proportional Symbols

Set **tmap** mode to interactive viewing:

```{r}
tmap_mode("view")
```

\
Map:

```{r}
tm_shape(sgpools_sf)+
tm_bubbles(col = "red",
           size = 1,
           border.col = "black",
           border.lwd = 1)
```

\
Make it proportional:

```{r}
tm_shape(sgpools_sf)+
tm_bubbles(
  col = "royalblue3",
           size = "Gp1Gp2 Winnings",
           border.col = "azure",
           border.lwd = 1)
```

\
Fill bubble colours by group and create facet plots:

```{r}
tm_shape(sgpools_sf) +
  tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1) +
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)
```

## Remember

\
to switch **tmap** viewer back to "plot" (static) mode at the end of each session!

```{r}
tmap_mode("plot")
```
