---
title: "Choropleths"
author: "farrahmf"
date: "25 February 2023"
date-modified: "`r Sys.Date()`"
execute: 
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

### Loading packages and importing datasets

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

This dataset records the number of Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling. (Source: [Singapore Department of Statistics](https://www.singstat.gov.sg/-/media/files/find_data/population/statistical_tables/respopagesextod2011to2020.ashx))

```{r}
popdata <- read_csv("aspatial/respopagesextod2011to2020.csv")
```

\
This will provide the geospatial data of our area of interest (Singapore), to enable mapping later on. Note that ".shp" files are multifile.

```{r}
mpsz <- st_read(dsn = "geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

\
Preparing a data table comprising variables *PA*, *SZ*, *YOUNG*, *ECONOMY ACTIVE*, *AGED*, *TOTAL* and *DEPENDENCY*, for the year 2020:

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:11]) +
           rowSums(.[13:15]))%>%
  mutate(`AGED`=rowSums(.[16:21])) %>% 
  mutate(`TOTAL`=rowSums(.[3:21])) %>%  
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
         `ECONOMY ACTIVE`, `AGED`, 
         `TOTAL`, `DEPENDENCY`)
```

\
The *PA* and *SZ* variables from **popdata2020** are intended to match the *PL_AREA_N* and *SUBZONE_N* variables from **mpsz**. As the former are in a mix of upper and lower case, while the latter are in upper case only, we need to convert the former to upper case only in order to join the two datasets.

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

\
Left-join **popdata2020** to **mpsz** to ensure the resulting table has complete set of geospatial data to enable mapping later on. Join on condition of matching SZ to SUBZONE_N. (All rows in **mpsz** would be retained, while only those rows in **popdata2020** that match rows in mpsz will be retained.)

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
write_rds(mpsz_pop2020, "rds/mpszpop2020.rds")
```

## Using *tmap*

The function qtm() will quickly draw a cartographic standard choropleth map:

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

\
*(Note: "view" as opposed to "mode" will provide an interactive version. Any subsequent mapping will adopt the last type used, i.e. "view" or "mode". Also, it won't work to simply replace "mode" with "view", other functions are needed (?).)*

\

The disadvantge of **qtm()** is that it makes aesthetics of individual layers harder to control. To draw a high quality cartographic choropleth map, **tmap**'s drawing elements should be used.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

#### Drawing a base map

**tm_shape()** : initiates the map

**tm_polygons()** : lays on the polygons defining the sub-areas of interest (in this case planning area sub-zones). The argument in the brackets will fill each polygon with a colour corresponding to the value associated with that variable for that sub-area. Values, colour scales and missing values are default. Wrapper for:

-   **tm_fill()** : arguments include **n** (number of classes), **style** (discrete and continuous gradient options) and most standard aesthetics

-   **tm_border()** : arguments include most standard aesthetics

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

```{r}
# the "jenks" method is somewhere between the "equal" and "quantile" methods
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "sd") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "kmeans") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "pretty") +
  tm_borders(alpha = 0.5)
```

\

Setting break points - start by looking at the descriptive statistics.

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

\
Based on the above, we can set break points at 0.60, 0.70, 0.80, and 0.90, with minimum and maximum of 0 and 1.00.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

\
Choosing colour schemes from the RColorBrewer palettes.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "BuPu") +
  tm_borders(alpha = 0.5)
```

\
Reverse the colour scheme:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "-BuPu") +
  tm_borders(alpha = 0.5)
```

Map layout components include various:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "BuPu", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("RIGHT", "BOTTOM"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

\
Map style options:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("cobalt")
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("natural")
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("bw")
```

\

Cartographic furniture: **tmap** also also provides arguments to draw other map furniture such as compass, scale bar and grid lines.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

 \
To reset to default style:

```{r}
tmap_style("white")
```

### Drawing multiple maps

By defining ***ncols*** in **tm_fill()**:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

\
Or by assigning multiple values to at least one of the aesthetic arguments:

```{r}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

\

Or by using **tm_facets()**:

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

\
Or by creating multiple stand-alone maps with **tmap_arrange()**:

```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### 

### **Mapping Spatial Object Meeting a Selection Criterion**

```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```
