---
title: "Hands-on Exercise 3"
author: "farrahmf"
---

### 1. Installing packages

```{r}
pacman::p_load(ggiraph, plotly, gganimate, DT, patchwork, tidyverse)
```

### 2. Importing data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

### 3. Tooltip

Using aesthetic **tooltip = ID** and **girafe()** to display the id of an element when the mouse hovers over it.

```{r}
p <- ggplot(data = exam_data,
            aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot"
  ) +
  scale_y_continuous(NULL, breaks = NULL)

girafe(ggobj = p,
       width_svg = 6,
       height_svg = 6*0.618)
```

Displaying customised information by creating a new object.

girafe(ggobj = p,

width_svg = 8,

height_svg = 8\*0.618

)

```{r}
exam_data$tooltip <- c(paste0(
  "Name = ", exam_data$ID,
  "\n Class = ", exam_data$CLASS
))

p <- ggplot(data = exam_data,
            aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot"
  ) +
  scale_y_continuous(NULL, breaks = NULL)

girafe(ggobj = p,
       width_svg = 8,
       height_svg = 8*0.618
)
```

Customise tooltop style using **opts_tooltip()** and **css** declarations.

```{r}
tooltip_css <- "background-color:DarkSlateGrey; font-family: Arial, Times, serif; color:LightYellow;"

p <- ggplot(data = exam_data,
            aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot"
  ) +
  scale_y_continuous(NULL, breaks = NULL)

girafe(ggobj = p,
       width_svg = 8,
       height_svg = 8*0.618,
       options = list(
         opts_tooltip(
           css = tooltip_css
         )
       )
)
```

Displaying statistics on tooltip.

```{r}
tooltip <- function(y, ymax, accuracy = .01) {
  mean <- scales::number(y, accuracy = accuracy)
  sem <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean, "+/-", sem)
}

gg_point <- ggplot(data = exam_data,
                   aes(x = RACE)) +
  stat_summary(aes(y = MATHS,
                   tooltip = after_stat(
                     tooltip(y, ymax))), 
               fun.data = "mean_se",
               geom = GeomInteractiveCol,
               fill = "#66CC99") +
  stat_summary(aes(y = MATHS),
               fun.data = mean_se,
               geom = "errorbar", width = 0.2, linewidth = 0.2)

girafe(ggobj = gg_point,
       width_svg = 8,
       height_svg = 8*0.618)
```

Highlighting elements associated with a **data_id**.

```{r}
p <- ggplot(data = exam_data,
            aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(data_id = CLASS),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot"
  ) +
  scale_y_continuous(NULL, breaks = NULL)

girafe(ggobj = p,
       width_svg = 8,
       height_svg = 8*0.618
)
```

As above, but inverting the highlight (i.e. lowlighting elements *not* associated with the data_id instead).

```{r}
p <- ggplot(data = exam_data,
            aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(data_id = CLASS),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot"
  ) +
  scale_y_continuous(NULL, breaks = NULL)

girafe(ggobj = p,
       width_svg = 8,
       height_svg = 8*0.618,
       options = list(
         opts_hover(css = "fill: #202020;"),
         opts_hover_inv(css = "opacity:0.2;")
       )
)
```

Using **onclick** to link to a website.

```{r}
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
                             "https://www.moe.gov.sg/schoolfinder?journey=Primary%20school", 
                             as.character(exam_data$ID))

p <- ggplot(data = exam_data,
            aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(onclick = onclick),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot"
  ) +
  scale_y_continuous(NULL, breaks = NULL)

girafe(ggobj = p,
       width_svg = 8,
       height_svg = 8*0.618)
```

Coordinated multiple views, using **ggiraph** and **patchwork**: when a data point on one plot is selected, the corresponding data point (i.e. sharing the same data ID) in another plot will be highlighted too.

```{r}
p1 <- ggplot(data = exam_data,
             aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(data_id = ID),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot"
  ) +
  coord_cartesian(xlim = c(0,100)) +
  scale_y_continuous(NULL, breaks = NULL)

p2 <- ggplot(data = exam_data,
             aes(x = ENGLISH)) +
  geom_dotplot_interactive(
    aes(data_id = ID),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot"
  ) +
  coord_cartesian(xlim = c(0,100)) +
  scale_y_continuous(NULL, breaks = NULL)

girafe(code = print(p1/p2),
       width_svg = 6,
       height_svg = 6,
       options = list(
         opts_hover(css = "fill: #202020;"),
         opts_hover_inv(css = "opacity:0.2")
       )
       )
```
