---
title: "Take-home Exercise 3"
author: "farrahmf"
date: "9 February 2023"
date-modified: "`r Sys.Date()`"
execute: 
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

In this exercise, we are tasked with building data visualisations using a dataset of HDB resale prices for sales registered since 1 January 2017 (source: [Data.gov.sg](https://data.gov.sg/dataset/resale-flat-prices?resource_id=f1765b54-a209-4718-8d38-a39237f502b3)).

## Loading packages and importing dataset

```{r}
pacman::p_load(ggstatsplot, tidyverse)

```

```{r}
resale_data <- read_csv("data/resale-flat-prices-based-on-registration-date-from-jan-2017-onwards.csv")
```

\[Resale price trends for each flat type since 1 Jan 17\]

```{r}
x <- "2017-01"
my_date <- as.Date(paste(x, "-01", sep=""))
my_date
class(my_date)
```

```{r}

resale_data1 <- resale_data %>%
  mutate(month = as.Date(paste(month, "-01", sep=""))) %>%
  mutate(category = flat_type) %>%
  mutate(category = ifelse(category == "1 ROOM", "1 to 2 Room", category)) %>%
  mutate(category = ifelse(category == "2 ROOM", "1 to 2 Room", category)) %>%
  mutate(category = ifelse(category == "3 ROOM", "3 to 4 Room", category)) %>%
  mutate(category = ifelse(category == "4 ROOM", "3 to 4 Room", category)) %>%
  mutate(category = ifelse(category == "5 ROOM", "5 Room or Other", category)) %>%
  mutate(category = ifelse(category == "EXECUTIVE", "5 Room or Other", category)) %>%
  mutate(category = ifelse(category == "MULTI-GENERATION", "5 Room or Other", category)) %>%
  group_by(month, category) %>%
  summarise(AvgRP = mean(resale_price))



## %>%
##  group_by(month, group) %>%
##   summarise(AvgRP = mean(resale_price))
```

```{r}
ggplot(data=resale_data1, aes(x=month, y=AvgRP, color=category)) +
  geom_line(aes(linetype=category))
```

\[Has the average size of resale flats declined year on year since 2017? (2023 excluded for too few observations)\]

```{r}
resale_data2 <- resale_data %>%
  mutate(year = str_replace(month, "-\\d\\d", "")) 
```

```{r}
pacman::p_load(ggstatsplot)
```

```{r}
set.seed(1234)

gghistostats(
  data = resale_data,
  x = floor_area_sqm,
  type = "bayes",
  test.value = 100,
  xlab = "Floor Area (sqm)"
)
```

```{r}
resale_data3 <- resale_data2 %>%
  filter(flat_type == "5 ROOM" | 
           flat_type == "EXECUTIVE" | flat_type == "MULTI-GENERATION")


```

```{r}
ggbetweenstats(
  data = resale_data3,
  x = year, 
  y = floor_area_sqm,
  type = "p",
  pairwise.comparisons = FALSE,
  messages = FALSE
)
```

\[difference in the floor area of the three large flat types\]

```{r}
ggbetweenstats(
  data = resale_data3,
  x = flat_type, 
  y = floor_area_sqm,
  type = "p",
  pairwise.comparisons = FALSE,
  messages = FALSE
)
```

\[difference retail price per square metre across three large flat types\]

```{r}
resale_data4 <- resale_data3 %>%
  mutate(psm = (resale_price/floor_area_sqm))
```

```{r}
ggbetweenstats(
  data = resale_data4,
  x = flat_type, 
  y = psm,
  type = "p",
  pairwise.comparisons = FALSE,
  messages = FALSE
)
```
